### 2021-12-13 documents/sections/verbs.tex
clarified jEpI reconstruction

The Pemongan languages and \kaxui point to \rc{jəpɨ} rather than \rc{epɨ}, although \kaxui \obj{johɨ} is very rare in contrast to more frequent \obj{o(o)hɨ}.
It only occurs in the third person form of the Progressive, meaning that the \obj{j} may be a reflex of third person \rc{i-}.
However, regular \obj{o}-initial \kaxui verbs have no third person prefix \pcref{tab:kaxprog}, while \obj{i-} occurs with C-initial \obj{to[mo]} \qu{to go}, suggesting that \obj{j} is indeed part of the root.
Putative \PC \rc{jəpɨ} would yield \rc{epɨ} by the two widespread sound changes of \rc{ə}-umlaut after \rc{j} and \rc{j}-loss \parencite{meira2010origin}, but one would expect evidence of a sequence \rc{jə} in \yawarana, \mapoyo, or \panare \parencite{gildea2010story}.
The \kaxui form \obj{ehɨ} would also not be accounted for, as \rc{ə} was not umlauted after \rc{j} in Parukotoan \parencite[501--502]{meira2010origin}.
%
\input{floats/kaxprog}%

>>>>

The Pemongan languages and \kaxui point to \rc{jəpɨ} rather than \rc{epɨ}, although \kaxui \obj{johɨ} is very rare in contrast to more frequent \obj{o(o)hɨ}.
It only occurs in the third person form of the Progressive, meaning that the \obj{j} may be a reflex of third person \rc{i-}.
However, regular \obj{o}-initial \kaxui verbs have no third person prefix \pcref{tab:kaxprog}, while \obj{i-} occurs with C-initial \obj{to[mo]} \qu{to go}, suggesting that \obj{j} is indeed part of the root.
Putative \PC \rc{jəpɨ} would yield \rc{epɨ} via the two widespread sound changes \phonl{\rc{ə}}{\rc{e}}{\rc{j}} and \phonl{\rc{j}}{∅}{\#} \parencite{meira2010origin}.
%Still, one would expect evidence of the sequence \rc{jə} in \yawarana, \mapoyo, or \panare, and the \kaxui form \obj{ehɨ} would also not be accounted for
The argument against this reconstruction is that \phonl{\rc{ə}}{\rc{e}}{\rc{j}} did not happen in \PPar, and only inconsistently in \PPP and \mapoyo-\yawarana \parencites[501--502]{meira2010origin}{gildea2010story}.
Thus, it does not explain \kaxui \obj{ehɨ}, nor the absence of the sequence \rc{jə} in \panare and \mapoyo-\yawarana.
%
\input{floats/kaxprog}%
